Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-14 paper.tex
Almost final

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}

@@

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities, such as expert disagreement and problem difficulty, that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

@@

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

The experts made updates on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days. The average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the data are very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

 The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

@@

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

\subsection{Discussion}
If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}). To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where 
\begin{eqnarray}
\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
\end{eqnarray}
This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
\begin{eqnarray*}
\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
\end{eqnarray*}
with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$. If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero. Since the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero. Therefore both the response and explanatory variables are approximately centered. This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule. If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter. This, however, reduces the interpretability of our model. Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form. One procedure to attain this is described in the beginning of Section \ref{realData}.

If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, must be chosen as the base-case to ensure that the probabilities will sum to one at any given time point. Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}). Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case. This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.

@@

To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Every grid point is used  40 times to generate a synthetic dataset. The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively. The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.

@@

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row presents the marginal effects of the three grid variables on the accuracy to which the SAC-procedure estimates the hidden process.  The second row is similar in nature but measures the accuracy in estimating the bias vector instead.  Since the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values. Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and perform better when $\beta$ is large.

@@

\subsection{Incoherent and Imbalanced Data}
\label{practicalMatters}
The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively. This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.  Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.

@@

\subsection{In- and Out-of-Sample Sharpness and Calibration}
\label{calibration}
A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model. The idea is to plot the probability forecasts against the observed empirical frequencies. Therefore any deviation from the diagonal line suggests poor calibration. A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend. In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot. Since the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.

>>>>

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Probability Aggregation in Time-Series}

@@

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item borrowing strength across the hierarchy to analyze the under- and overconfidence in different groups of experts,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
\end{itemize}

@@

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Given that a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

The experts made updates on a very infrequent basis: the average number of forecasts per expert was around 0.017 forecasts per day, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

@@

 The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Given that each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

@@

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

\subsection{Discussion}
If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}). To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where 
\begin{eqnarray}
\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
\end{eqnarray}
This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
\begin{eqnarray*}
\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
\end{eqnarray*}
with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$. If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero. Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero. Therefore both the response and explanatory variables are approximately centered. This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule. If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter. This, however, reduces the interpretability of our model. Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form. One procedure to attain this is described in the beginning of Section \ref{realData}.

If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, must be chosen as the base-case to ensure that the probabilities will sum to one at any given time point. Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}). Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case. This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.

@@

To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Every grid point is used  40 times to generate a synthetic dataset. The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.

%$|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively. The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.

@@

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row presents the marginal effects of the three grid variables on the accuracy to which the SAC-procedure estimates the hidden process.  The second row is similar in nature but measures the accuracy in estimating the bias vector instead.  Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values. Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and present a trade-off between estimating $X_{t,k}$ and $\boldsymbol{b}$ under different values of $\beta$. This trade-off arises from \ville{WHY DOES THIS HAPPEN?}

@@

\subsection{Incoherent and Imbalanced Data}
\label{practicalMatters}
The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively. This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Given that this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.  Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.

@@

\subsection{In- and Out-of-Sample Sharpness and Calibration}
\label{calibration}
A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model. The idea is to plot the probability forecasts against the observed empirical frequencies. Therefore any deviation from the diagonal line suggests poor calibration. A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend. To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot. Given that the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.
